---
title: "Sales forecasting for the European drug store Rossman"
# Set the formatting options
format:
  pdf:
    pdf-engine: xelatex
# Include a Table Of Contents
    toc: true
# Include a List Of Figures
    lof: false
# Include a List Of Tables
    lot: false
# Number each headed section
    number-sections: true
# Set the main font size and font
    fontsize: 11pt
    mainfont: Calibri
# Adjust the borders to decrease or increase the useable page space
    geometry:
      - top = 15mm
      - bottom = 20mm
      - left = 20mm
      - right = 20mm
# Adjust the headers and footers
    header-includes:
# Set the packages to be used by LaTex
      - \usepackage{placeins}
      - \usepackage{fancyhdr}
      - \usepackage{lastpage}
#  Set the style and what goes in the header and footer of the main and all other pages.
      - \pagestyle{fancy}
      - \thispagestyle{fancy}
      - \fancyhead[R]{Student ID| 11600159}
      - \fancyhead[L]{DATA70002 | Understanding Data and its Environment}
      - \renewcommand{\headrulewidth}{0.02pt}
      - \fancypagestyle{plain}{\fancyhead[R]{Student ID| 11600159}\fancyhead[L]{DATA70002 | Understanding Data and its Environment}\fancyfoot[C]{Page \thepage\ of \pageref{LastPage}}}
# Set the page number to be "Page n of np"
      - \fancyfoot[C]{Page \thepage\ of \pageref{LastPage}}

# CSL and bibliography file to be used
bibliography: DATA70002_refs.bib
csl: "UoL_Harvard.csl"
---

```{=html}
<style>
figcaption {
    text-align: center;
}
</style>
```

```{r , echo=FALSE, include=FALSE, eval=FALSE, message=FALSE}
# Load the required libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
```

```{r , echo=FALSE, include=FALSE, eval=FALSE, message=FALSE}
# Load the data
storedata <- read.csv("Data/clean_store_data.csv")
test_data <- read.csv("Data/clean_test_data.csv")
traindata <- read.csv("Data/clean_train_data.csv")
```


\newpage

# Introduction

# Methodology

### Store data cleaning
```{r}
# Convert storedata$Assortment
# a(basic) to "1"
# b(extra) to "2"
# c(extended) to "3"
table(storedata$Assortment)
storedata$Assortment <- recode(storedata$Assortment, "a" = 1, "b" = 2, "c" = 3)
table(storedata$Assortment)
```

```{r}
# impute the 3 missing values with a mean distance
mean(storedata$CompetitionDistance)
storedata$CompetitionDistance[is.na(storedata$CompetitionDistance)] <- mean(storedata$CompetitionDistance)
# Then rescale storedata$CompetitionDistance into deciles
# 1st into quantiles so outliers are more irrelevant, then into deciles
# Rescale 'value' column into deciles
storedata$CompetitionDistance_dec <- cut(storedata$CompetitionDistance, breaks = quantile(storedata$CompetitionDistance, probs = seq(0, 1, by = 0.1)), 
                 include.lowest = TRUE, labels = FALSE)

# Now move the new column so that it sits next to the original column
storedata <- storedata |> 
  relocate(CompetitionDistance_dec, .after = CompetitionDistance)
```




Review the available data and describe it in terms of its variables, quality, and relevance to the sales forecasting
































Link data sets together as appropriate

Pre-process the data as appropriate for further analytics, for example, you may want to encode any categorical data, create new variables, identify how many missing values there are and deal with them appropriately, etc.

Identify the key factors affecting sales, for example, you may want to check whether competition and promotions have an impact on sales, and how public holidays cause sales fluctuations.

Build a forecasting model (which can be a linear regression model, a neural network model or something else) using the variables you identified. Please make sure to justify the choice of your modelling approach.

Use the Root Mean Square Percentage Error (RMSPE) to forecast accuracy

# Results

Interpret key results, assumptions and limitations of your analysis.

# Conclusion

## Limitations

## Implications

## Recommendations

# References
